A Partitioning Algorithm for Maximum Common Subgraph Problems

### 2017-08-09 poster/poster.tex
First draft of poster

{
    \colorlet{blockbodybgcolor}{uofgcobalt}
    \colorlet{blocktitlebgcolor}{uofgcobalt}
    \block[bodyverticalshift=0cm, bodyinnersep=3mm]{}{
        \centering\begin{minipage}{0.94\textwidth}
            We introduce a new branch and bound algorithm for the \textbf{maximum common
            subgraph} and \textbf{maximum common connected subgraph} problems which is based
            around vertex labelling and partitioning. Our method resembles
            a constraint programming approach, but uses a novel compact
            domain store to reduce
            the memory and computation requirements during search.  Experiments
            show a speedup of more than an order of magnitude over the state of the
            art, and demonstrate that we can operate on much larger graphs without
            running out of memory.
%            We show how to generate \textbf{really hard} random instances for \textbf{subgraph
%            isomorphism} problems. For the non-induced variant, we predict and observe a phase
%            transition between satisfiable and unsatisfiable instances, with a corresponding
%            complexity peak seen in \textbf{three different solvers}. For the induced variant, much
%            richer behaviour is observed, and \textbf{constrainedness} gives a better measure of
%            difficulty than does proximity to a phase transition. We also discuss \textbf{variable
%            and value ordering} heuristics, and their relationship to the \textbf{expected number of
%            solutions}.
        \end{minipage}
    }
}

@@

\block{The Maximum Common Subgraph Problem}{
\begin{wrapfigure}[7]{r}{0.46\linewidth}
    \begin{center}\begin{tikzpicture}[scale=1.4]%{{{
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nc) at (1, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nd) at (3, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (Ne) at (3, -2)
        {\vphantom{0}};

@@

        \draw [ledge] (N1) -- (N2);
        \draw [edge] (N1) -- (N3);
        \draw [ledge] (N1) -- (N4);
        \draw [ledge] (N2) -- (N4);
        \draw [edge] (N3) -- (N5);
        \draw [edge] (N3) -- (N6);
        \draw [ledge] (N4) -- (N6);
        \draw [edge] (N5) -- (N6);
        \draw [ledge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}

@@

The \textbf{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two given graphs simultaneously.
Because graphs are widely used to model real-world phenomena, maximum common
subgraph problems have arisen in molecular science (where graphs often represent
molecules)
and also in other domains including malware detection, source code analysis, and computer vision.

@@

Maximum common subgraph problems are NP-hard, and remain challenging
computationally. Recent practical progress has been made by using constraint
programming and
mathematical programming, by reducing
to the maximum clique problem, and by
adapting subgraph isomorphism algorithms. Some
special cases also have practical polynomial time algorithms.

@@

We consider the \textbf{maximum common induced subgraph} problem, in
which the objective is to find a graph with as many vertices as possible which
is an induced subgraph of each of two input graphs.  We introduce
a new branch and bound algorithm which exploits special properties of the
problem to allow a much faster exploration of the search space, whilst
retaining the filtering and bounding benefits of the constraint programming
approach. 
}

\block{What goes here?}{

%If we randomly generate a pattern graph with $p$ vertices and density $d_p$, and a target graph
%with $t$ vertices and density $d_t$, the \textbf{expected number of solutions} to the non-induced
%isomorphism problem is \[
%    \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot {d_t}^{d_p \cdot
%    \binom{p}{2}} } \textnormal{.} \]

Walkthrough?

@@

Parallel version??

%By considering when $\mathsf{\langle Sol \rangle = 1}$, we can \textbf{estimate the location} of the
%satisfiable / unsatisfiable phase transition. (This is not entirely accurate due to variance,
%particularly when the pattern is very sparse or very dense).

@@

Extensions to labelled and connected cases???

%We can also use this formula to recover \textbf{variable} and \textbf{value ordering heuristics}: we
%should make branching decisions which maximise the expected number of solutions during search. This
%tells us to prefer small domains and pattern vertices of high degree, but target vertices of low
%degree, which matches the empirically-determined heuristics used in state of the art solvers.

@@

@@

\block{Horse Race}{
We implemented \McSplit in C++. We compare against the best constraint
programming implementations 
(CP-FC in the unlabelled cases, and CP-MAC
in the labelled cases, using both branching and filtering for connected
subgraphs), the encodings, and the
$k{\downarrow}$ algorithm (which only supports
unlabelled, undirected, unconnected instances).  Each of these comparator
programs is an optimised, dedicated implementation and does not use a
general-purpose constraint programming toolkit.

Our first set of experiments uses a database of randomly-generated maximum
common subgraph instances.  For
unlabelled instances, we selected the first ten instances from each family
whose members have no more than 50 vertices, for a total of 4,100 instances.
For labelled instances, we selected the first ten instances from every family,
for a total of 8,140 instances with up to 100 vertices;
we use the labelling scheme in which the
number of distinct vertex labels and the number of distinct edge labels is
approximately equal to 33 percent of the number of vertices in each graph.

    \vspace*{1em}

    \begin{center}
        \small
        \input{gen-graph-plain-cumulative}
        %\input{gen-graph-phase-transition}
    \end{center}

%    We can also look at what happens if we vary both the pattern and the target density, for
%    different orders of pattern. The top row shows satisfiability---the black lines show where
%    $\mathsf{\langle Sol \rangle = 1}$. The bottom row shows the search cost for one of the three
%    algorithms we consider in the paper. As expected, the ``really hard'' instances are near the
%    phase transition.

    \vspace*{0.9em}

    \begin{center}
        \small
        %\input{gen-graph-non-induced}
    \end{center}
}

\block{Is this Just a Tweaked Version of CP-FC?}{
\begin{wrapfigure}[24]{r}{0.64\linewidth}\begin{center}
    %\small\input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
\end{center}\end{wrapfigure}

    Yes.

    We show that here with a nice heatmap and a discussion of the soft all-diff propagator,
    maybe with a bipartite graph figure.

%    For induced isomorphisms, the behaviour is much richer. We can derive an expected number of
%    solutions of \[
%        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot
%                    {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} } \textnormal{.} \]
%
%    \bigskip
%
%    Again this predicts a sharp phase transition. When the pattern is small, we see hard instances
%    near the phase transition, and easy instances elsewhere. However, when the pattern is larger,
%    the central unsatisfiable region \textbf{remains very hard}, even though it is no longer near a
%    phase transition.
%
%    \bigskip
%
%    This is not just a weakness of current subgraph isomorphism algorithms: the region is also hard
%    for pseudo-boolean and boolean satisfiability solvers, and under reduction to the clique
%    problem.
%
%    \bigskip
%
%    The central hard region \emph{is} predicted correctly by \textbf{constrainedness}, as we
%    show in the third row of plots. Although far from a phase transition, this region is only
%    slightly overconstrained.
%
%    \bigskip
%
%    Looking to maximise $\mathsf{\langle Sol \rangle}$ derives existing variable and value ordering
%    heuristics, in the case that both the pattern and target graphs are sparse. The formula suggests
%    that algorithms should swap heuristics when the pattern and target are dense---empirically, this
%    does indeed give an improvement. However, maximising the expected number of solutions is
%    \textbf{not enough to select the best heuristic} in two eighths of the search space (and
%    constrainedness does not help).
}

\block{See the Paper For\ldots}{
    \begin{itemize}
        \item ~ How the algorithm can be (straightforwardly extended) to handle vertex and edge labels
        \item ~ Experimental results for the labelled and connected cases
        \item ~ ...???
    \end{itemize}
}

\block{Future Work}{
    \begin{itemize}
        \item ~ Better branching heuristics
        \item ~ Branching on vertices in both graphs, not just the first one
        \item ~ Applying the partitioning data structure to other problems
    \end{itemize}
}

>>>>

\begin{columns}
\column{1}
{
    \colorlet{blockbodybgcolor}{uofgcobalt}
    \colorlet{blocktitlebgcolor}{uofgcobalt}
    \block[bodyverticalshift=0.5cm, bodyinnersep=3mm]{}{
        \centering\begin{minipage}{0.94\textwidth}
            \Large We introduce a new branch and bound algorithm for the \textbf{maximum common
            subgraph} and \textbf{maximum common connected subgraph} problems which is based
            around vertex labelling and partitioning. Our method resembles
            a constraint programming approach, but uses a novel compact
            domain store to reduce
            the memory and computation requirements.  Experiments
            show a speedup of more than an order of magnitude over the state of the
            art, and demonstrate that we can operate on much larger graphs without
            running out of memory.
%            We show how to generate \textbf{really hard} random instances for \textbf{subgraph
%            isomorphism} problems. For the non-induced variant, we predict and observe a phase
%            transition between satisfiable and unsatisfiable instances, with a corresponding
%            complexity peak seen in \textbf{three different solvers}. For the induced variant, much
%            richer behaviour is observed, and \textbf{constrainedness} gives a better measure of
%            difficulty than does proximity to a phase transition. We also discuss \textbf{variable
%            and value ordering} heuristics, and their relationship to the \textbf{expected number of
%            solutions}.
        \end{minipage}
        \vspace*{0.4cm}
    }
}
\end{columns}

@@

\block{The Maximum Common Subgraph Problem}{
\begin{wrapfigure}[7]{r}{0.48\linewidth}
    \begin{flushright}\begin{tikzpicture}[scale=1.6]%{{{
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nc) at (1, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nd) at (3, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (Ne) at (3, -2)
        {\vphantom{0}};

@@

        \draw [ledge] (N1) -- (N2);
        \draw [edge] (N1) -- (N3);
        \draw [ledge] (N1) -- (N4);
        \draw [ledge] (N2) -- (N4);
        \draw [edge] (N3) -- (N5);
        \draw [edge] (N3) -- (N6);
        \draw [ledge] (N4) -- (N6);
        \draw [edge] (N5) -- (N6);
        \draw [ledge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{flushright}

@@

The \textbf{maximum common subgraph} family of problems involves finding a large
graph which is isomorphic to subgraphs of two input graphs.
Such problems have arisen in molecular science---where graphs often represent
molecules---malware detection, source code analysis, and computer vision.

@@

    Maximum common subgraph problems are \mbox{NP-hard}, and remain challenging
computationally. Recent practical progress has been made by using constraint
programming and
mathematical programming, by reducing
to the maximum clique problem, and by
adapting subgraph isomorphism algorithms. Some
special cases also have practical polynomial time algorithms.

@@

We consider the \textbf{maximum common induced subgraph} problem, in
which the objective is to find a graph with as many vertices as possible which
is an induced subgraph of each of two input graphs.  The diagram shows a simple instance
with an optimal solution highlighted.
}

\block{The \McSplit\ Algorithm}{

We introduce a branch and bound algorithm which exploits special properties of the
problem to allow a much faster exploration of the search space, whilst
retaining the filtering and bounding benefits of a constraint programming
approach.

@@

Using depth-first search, the \McSplit\ algorithm builds up an injective mapping from a subset
of vertices in the pattern graph to a subset of vertices in the target graph.  
Remaining vertices are partitioned into \textbf{label classes} according to the set of mapped
vertices to which they are adjacent; each pattern-graph vertex may only be mapped to target-graph
vertices in the same label class.  We thus avoid
the need for a conventional domain store in which the possible mappings for each
vertex are stored explicitly.

@@

Straightforward extensions can be made to the algorithm to handle edge- and vertex-labelled graphs,
and directed graphs.  We can also solve the problem where the subgraph must be connected,
by branching only on vertices adjacent to at least one currently-mapped vertex.
}

\block{Relationship to Prior Constraint Programming Encodings}{

    Our experimental results suggest that \McSplit\ has broadly
    similar performance trends to the constraint programming,
    forward-checking (CP-FC) algorithm of Ndiaye and Solnon
    (2011), but with much lower constant factors and memory
    usage.  In the figure below, we plot the number
    of recursive calls made by our algorithm versus the number
    made by CP-FC, for unlabelled, unconnected
    problem instances. (Rather than a simple scatter plot,
    darker colours are used to indicate a higher density of points around
    a location.) We see a close correlation: \McSplit\ typically
    does slightly less work, and sometimes does more, but instances with
    more than one order of magnitude difference in
    search tree size are rare.

    Why is this?  The key
    observation is that \McSplit\ may be considered to be a different version of the CP-FC algorithm,
    using an unconventional domain store and more efficient filtering algorithms.
    We explore this relationship in the paper.

\begin{center}
    \small\input{gen-graph-plain-james-versus-cp-fc-nodes-scatter}
\end{center}
}

\block{Experiments}{
We compare our implementation of \McSplit\ against the best existing constraint
programming implementations, clique encodings, and the
$k{\downarrow}$ algorithm of McCreesh et al. (2017).  Each of these comparator
programs is an optimised, dedicated implementation and does not use a
general-purpose constraint programming toolkit.

For the unlabelled problem, we used 4,100 instances from a
database of randomly-generated maximum common subgraph instances.  For
the labelled variant, we used 8,140 instances from the same database.  \McSplit\ is
more than an order of magnitude faster than its nearest competitor on unlabelled instances.
On labelled instances, the clique encoding remains the single best solver.

    \vspace*{2em}

    \begin{center}
        \small
        \input{gen-graph-plain-cumulative}
    \end{center}

    \vspace*{2em}

    \begin{center}
        \small
        \input{gen-graph-33ved-cumulative}
    \end{center}

    \vspace*{3em}

We also ran the algorithms on a set of 5,725 larger instances used in recent
studies of subgraph isomorphism and maximum common subgraph; a modified version
of \McSplit\ is the single best solver on this set of instances.

    \vspace*{3em}

@@

\block{See the Paper For\ldots}{
    \begin{itemize}
        \item ~ A walk-through of the algorithm
        \item ~ How to extend the algorithm to handle vertex and edge labels
        \item ~ Additional experimental results
    \end{itemize}
}

@@

\block{Future Work}{
    \begin{itemize}
        \item ~ Better branching heuristics
        \item ~ Branching on vertices in both graphs
        \item ~ Applying the partitioning data structure to other problems
    \end{itemize}
    \vspace*{0.25em}
}
